Motion Planning and Robust Fault-Tolerant Control of Hybrid UAVs and Biped Robots Team System for Search and Rescue Usage

### 2022-11-24 paper/paper.tex
small revise paper

\subsection{motion planning of flying of UAV}
A dynamic model about how the UAV moves in the physical world is given first. By Newton-Euler equation, the dynamic model of UAV can be formulated as \cite{sabatino2015quadrotor}:
\begin{equation} \label{eq:uav} 
    \begin{split}
        \begin{bmatrix}
            f_u \\ \tau_u
        \end{bmatrix}&=\begin{bmatrix}
            mI & 0 \\ 0 & J
        \end{bmatrix}\begin{bmatrix}
            \ddot{X} \\ \ddot{\Theta}
        \end{bmatrix}+\begin{bmatrix}
            0 \\ \dot{\Theta}\times(J\dot{\Theta})
        \end{bmatrix}+\begin{bmatrix}
            f_g \\ 0
        \end{bmatrix}
        \\
        &+\begin{bmatrix}
            K_F & 0 \\
            0 & K_\tau
        \end{bmatrix}\begin{bmatrix}
            \dot{X} \\ \dot{\Theta}
        \end{bmatrix}
        % form 2
        % &R(\Theta)=\begin{bmatrix*}[l]
        %     \cos\theta\cos\phi & \sin\phi\sin\theta\cos\psi-\cos\phi\sin\psi & R_1 \\
        %     \cos\theta\sin\phi & R_2 & R_3 \\
        %     -\sin\theta & \sin\phi\cos\theta & R_4
        % \end{bmatrix*}
        % \\
        % &R_1=\cos\phi\sin\theta\cos\psi+\sin\phi\sin\psi
        % \\
        % &R_2=\sin\phi\sin\theta\sin\psi+\cos\phi\cos\psi
        % \\
        % &R_3=\cos\phi\sin\theta\sin\psi-\sin\phi\cos\psi
        % \\
        % &R_4=\cos\phi\cos\theta
    \end{split}
\end{equation}
where $J=\begin{bmatrix}
    J_{x} & 0 & 0 \\
    0 & J_{y} & 0 \\
    0 & 0 & J_{z}
\end{bmatrix}$, $K_F=\begin{bmatrix}
    K_{x} & 0 & 0 \\
    0 & K_{y} & 0 \\
    0 & 0 & K_{z}
\end{bmatrix}$, $K_\tau=\begin{bmatrix}
    K_{\tau_x} & 0 & 0 \\
    0 & K_{\tau_y} & 0 \\
    0 & 0 & K_{\tau_z}
\end{bmatrix}$, $f_u =\begin{bmatrix}
    f_x \\ f_y \\ f_z
\end{bmatrix} = R(\Theta)\begin{bmatrix}
    0 \\ 0 \\ F
\end{bmatrix}$, $\tau_u=\begin{bmatrix}
    \tau_x \\ \tau_y \\ \tau_z
\end{bmatrix}$, $X=\begin{bmatrix}
    x \\ y \\ z
\end{bmatrix}$, $\Theta=\begin{bmatrix}
    \phi \\ \theta \\ \psi
\end{bmatrix}$, $f_g=\begin{bmatrix}
    0 \\ 0 \\ -mg
\end{bmatrix}$, $R(\Theta)=R_z(\psi)R_y(\theta)R_x(\phi)$, $R_z(\psi)=\begin{bmatrix} % form 1
    \cos\psi & -\sin\psi & 0 \\
    \sin\psi & \cos\psi & 0 \\
    0 & 0 & 1
\end{bmatrix}$, $R_y(\theta)=\begin{bmatrix}
    \cos\theta & 0 & \sin\theta \\
    0 & 1 & 0 \\
    -\sin\theta & 0 & \cos\theta
\end{bmatrix}$, $R_x(\phi)=\begin{bmatrix}
    1 & 0 & 0 \\
    0 & \cos\phi & -\sin\phi \\
    0 & \sin\phi & \cos\phi
\end{bmatrix}$.
% where $y'$ denotes the $y$ axis after rotating the $z$ axis and $x''$ denotes the $x$ axis after rotating the $y'$ axis
$g$ is the gravity acceleration, $m$ and $J$ are the mass and inertia matrix of UAV, respectively, $\tau_u$ and $F$ are the total torque and force acting on UAV, respectively, $\Theta$ is the Euler angles in body frame, $X$ is the postion of center of mass (CoM) in inertial frame, $K_\tau$ and $K_F$ are the aerodynamic damping coefficients, and $R(\Theta)$ is the intrinsic rotation matrix from body frame to inertial frame. This model treats the UAV as a mass point and can control the force in the $z$ direction and the torque in the $x$, $y$ and $z$ direction. Hence, the reference trajectory of UAV $r(t)=[x_r(t),y_r(t),z_r(t),\phi_r(t),\theta_r(t),\psi_r(t)]^T\in\mathbb{R}^{6}$, where the subscript $r$ denotes the reference.

@@

\end{bmatrix}^T\in\mathbb{R}^{(w+1)n}$, $A_i\in\mathbb{R}^{{(w+1)n}\times{(w+1)n}}$, $v_i(t)$ is the model error, $C_i=\begin{bmatrix}

@@

    \end{bmatrix}\otimes I_{n_i}$, and $w$ is the window size of smoothing signal model with $n_1=n$ and $n_2 = o$. Substituting (\ref{eq:smooth model}) into (\ref{eq:error}), we get the following augmented tracking error system of an agent:

@@

    \end{bmatrix} = 3n + (w-1)n + (w-1)o, \forall z \in eig(\bar{A})$.

@@

Although the control gain in (\ref{eq:u_fb}) and observer gain in (\ref{eq:e_bar}) for each agent in URTS can already be found through the previous steps, the calculation speed of solving the matrix inequality (\ref{BMI1}) and the online calculation speed of controller and observer can be further improved by reducing the dimensionality. Observing the matrices $A,B,C,B_2$ in the linearized system (\ref{eq:error}), it can be further split into $n$ subsystems for each agent if the matrices $C,B_2$ in the output equation (\ref{eq:agent output}) have the same form to $A,B$ and $l=l_0n, o=o_0n$, i.e., $C = C_0\otimes I_n,B_2 = B_{2,0}\otimes I_n$ where $C_0\in\mathbb{R}^{l_0\times 3},B_{2,0}\in\mathbb{R}^{l_0\times o_0}$. Let us decompose the error $\pmb{e}(t)=\sum_{i=1}^{n} \pmb{e}_i(t)\otimes\mathbf{e}_i$, the control $u_{fb,i}(t)=\sum_{i=1}^{n} {u}_{fb,i}(t)\otimes\mathbf{e}_i$, the acuaor fault $f_{1,i}(t)=\sum_{i=1}^{n} {f}_{1,i}(t)\otimes\mathbf{e}_i$, the output $y(t)=\sum_{i=1}^{n} {y}_i(t)\otimes\mathbf{e}_i$, and the sensor fault $f_{2,i}(t)=\sum_{i=1}^{n} {f}_{2,i}(t)\otimes\mathbf{e}_i$ where $\pmb{e}_i(t)\in\mathbb{R}^3$, $u_{fb,i}(t)\in\mathbb{R}^1$, $f_{1,i}(t)\in\mathbb{R}^1$, $y_i(t)\in\mathbb{R}^{l_0}$, $f_{2,i}(t)\in\mathbb{R}^{o_0}$ and $\mathbf{e}_i$ is standard unit column vectors in $\mathbb{R}^n$, we get the $n$ subsystems:
\begin{equation} \label{eq:linear subsys}
    \begin{split}
        & \dot{\pmb{e}}_i(t)=A_0\pmb{e}_i(t)+B_0(u_{fb,i}(t)+f_{1,i}(t)) \\
        & {y}_i(t)=C_0\pmb{e}_i(t)+B_{2,0}f_{2,i}(t)   
    \end{split}
\end{equation}
\begin{remark}
    If the linearized system (\ref{eq:error}) can be splited into $n$ subsystems, this means that $\pmb{e}_i(t)$, the PID error of each state variable $x(t)$ of each agent in (\ref{eq:agent}), can be measured independently via sensors to obtain the independent outputs $y_i(t)$. In actual systems, this is usually done.
\end{remark}

By Theorem \ref{theorem1} again, the form shows that we can find the control gain $K_i\in\mathbb{R}^{1\times s}$ and observer gain $L_i\in\mathbb{R}^{s\times l_0}$ of the $i$th subsystem (\ref{eq:linear subsys}) that achieve the $H_\infty$ observer-based tracking control performance with a prescribed attenuation level $\rho_i$, where $s=3+(w-1)+(w-1)o_0$. The origin control gain $K$ of the origin agent system can be reconstructed by $K = \begin{bmatrix}
    \mathit{k}_1 & \mathit{k}_2 & ... & \mathit{k}_n
\end{bmatrix}^T, \mathit{k}_i=K_i^T\otimes\mathbf{e}_i$. The origin observer gain $L$ can be reconstructed in the same way.

@@

\textit{Agents:}\begin{enumerate}
    \item system parameters:
    Initial value $\tilde{x}(0) = [\pmb{e}(0)^T, F_1(0)^T\mathbin{,} F_2(0)^T, \tilde{e}(0)]^T = [[0.1X, ..., 0.1X]^T, 0, 0, 0]^T$ where $X \sim \mathcal{N}(0, 1)$. $C = C_0 \otimes I_n$ where $C_0 = I_3$.
    \item designed parameters: $\rho^*=30$. Actuator and sensor window size of smoothing signal model are $3$ and $4$, respectively.
\end{enumerate}

@@

    \item actuator coupling disturbance in (\ref{eq:UAV couple}):
    \\$c(t) = \sum_{k = 1, k \neq i}^{N_T}D_{i, 1, k}(x_{i, 1}(t))x_{k, 1}(t)\in\mathbb{R}^6$ where $D_{i, 1, k}(x_{i, 1}(t)) = diag(x_{i, 1, 1}(t)\mathbin{,}\dots\mathbin{,}x_{i, 1, 6}(t))$ with $x_{i, 1}(t) = [x_{i, 1, 1}(t)\mathbin{,}\dots\mathbin{,}x_{i, 1, 6}(t)]^T$.
    \item actuator non-coupling disturbance:
    \\ $d_1(t) = [100\sin(3t)\mathbin{,} ...\mathbin{,} 10\sin(3t)]^T\in\mathbb{R}^6$.
    \item sensor fault: $f_2(t)$ is set as a smoothed square wave as shown in Fig. \ref{fig:UAV, fa}.
\end{enumerate}

>>>>

\subsection{motion planning of flying of UAV}
A dynamic model about how the UAV moves in the physical world is given first. By Newton-Euler equation, the dynamic model of UAV can be formulated as \cite{sabatino2015quadrotor}:
\begin{equation} \label{eq:uav} 
    \begin{split}
        \begin{bmatrix}
            f_u \\ \tau_u
        \end{bmatrix}&=\begin{bmatrix}
            mI & 0 \\ 0 & J
        \end{bmatrix}\begin{bmatrix}
            \ddot{X} \\ \ddot{\Theta}
        \end{bmatrix}+\begin{bmatrix}
            0 \\ \dot{\Theta}\times(J\dot{\Theta})
        \end{bmatrix}+\begin{bmatrix}
            f_g \\ 0
        \end{bmatrix}
        \\
        &+\begin{bmatrix}
            K_F & 0 \\
            0 & K_\tau
        \end{bmatrix}\begin{bmatrix}
            \dot{X} \\ \dot{\Theta}
        \end{bmatrix}
        % Form 2 of rotation matrix R(\Theta)
        % &R(\Theta)=\begin{bmatrix*}[l]
        %     \cos\theta\cos\phi & \sin\phi\sin\theta\cos\psi-\cos\phi\sin\psi & R_1 \\
        %     \cos\theta\sin\phi & R_2 & R_3 \\
        %     -\sin\theta & \sin\phi\cos\theta & R_4
        % \end{bmatrix*}
        % \\
        % &R_1=\cos\phi\sin\theta\cos\psi+\sin\phi\sin\psi
        % \\
        % &R_2=\sin\phi\sin\theta\sin\psi+\cos\phi\cos\psi
        % \\
        % &R_3=\cos\phi\sin\theta\sin\psi-\sin\phi\cos\psi
        % \\
        % &R_4=\cos\phi\cos\theta
    \end{split}
\end{equation}
where $J=\begin{bmatrix}
    J_{x} & 0 & 0 \\
    0 & J_{y} & 0 \\
    0 & 0 & J_{z}
\end{bmatrix}$, $K_F=\begin{bmatrix}
    K_{x} & 0 & 0 \\
    0 & K_{y} & 0 \\
    0 & 0 & K_{z}
\end{bmatrix}$, $K_\tau=\begin{bmatrix}
    K_{\tau_x} & 0 & 0 \\
    0 & K_{\tau_y} & 0 \\
    0 & 0 & K_{\tau_z}
\end{bmatrix}$, $f_u =\begin{bmatrix}
    f_x \\ f_y \\ f_z
\end{bmatrix} = R(\Theta)\begin{bmatrix}
    0 \\ 0 \\ F
\end{bmatrix}$, $\tau_u=\begin{bmatrix}
    \tau_x \\ \tau_y \\ \tau_z
\end{bmatrix}$, $X=\begin{bmatrix}
    x \\ y \\ z
\end{bmatrix}$, $\Theta=\begin{bmatrix}
    \phi \\ \theta \\ \psi
\end{bmatrix}$, $f_g=\begin{bmatrix}
    0 \\ 0 \\ -mg
\end{bmatrix}$, $R(\Theta)=R_z(\psi)R_y(\theta)R_x(\phi)$, $R_z(\psi)=\begin{bmatrix} % form 1
    \cos\psi & -\sin\psi & 0 \\
    \sin\psi & \cos\psi & 0 \\
    0 & 0 & 1
\end{bmatrix}$, $R_y(\theta)=\begin{bmatrix}
    \cos\theta & 0 & \sin\theta \\
    0 & 1 & 0 \\
    -\sin\theta & 0 & \cos\theta
\end{bmatrix}$, $R_x(\phi)=\begin{bmatrix}
    1 & 0 & 0 \\
    0 & \cos\phi & -\sin\phi \\
    0 & \sin\phi & \cos\phi
\end{bmatrix}$.
% where $y'$ denotes the $y$ axis after rotating the $z$ axis and $x''$ denotes the $x$ axis after rotating the $y'$ axis
$g$ is the gravity acceleration, $m$ and $J$ are the mass and inertia matrix of UAV, respectively, $\tau_u$ and $F$ are the total torque and force acting on UAV, respectively, $\Theta$ is the Euler angles in body frame, $X$ is the postion of center of mass (CoM) in inertial frame, $K_\tau$ and $K_F$ are the aerodynamic damping coefficients, and $R(\Theta)$ is the intrinsic rotation matrix from body frame to inertial frame. This model treats the UAV as a mass point and can control the force in the $z$ direction and the torque in the $x$, $y$ and $z$ direction. Hence, the reference trajectory of UAV $r(t)=[x_r(t),y_r(t),z_r(t),\phi_r(t),\theta_r(t),\psi_r(t)]^T\in\mathbb{R}^{6}$, where the subscript $r$ denotes the reference.

@@

\end{bmatrix}^T\in\mathbb{R}^{(w_i+1)n_i}$, $A_i\in\mathbb{R}^{{(w_i+1)n_i}\times{(w_i+1)n_i}}$, $v_i(t)$ is the model error, $C_i=\begin{bmatrix}

@@

    \end{bmatrix}\otimes I_{n_i}$, and $w_i$ is the window size of smoothing signal model with $n_1=n$ and $n_2 = o$. Substituting (\ref{eq:smooth model}) into (\ref{eq:error}), we get the following augmented tracking error system of an agent:

@@

    \end{bmatrix} = 3n + (w_1+1)n + (w_2+1)o, \forall z \in eig(\bar{A})$.

@@

Although the control gain in (\ref{eq:u_fb}) and observer gain in (\ref{eq:e_bar}) for each agent in URTS can already be found through the previous steps, the calculation speed of solving the matrix inequality (\ref{BMI1}) and the online calculation speed of controller and observer can be further improved by reducing the dimensionality. Observing the matrices $A,B,C,B_2$ in the linearized system (\ref{eq:error}), it can be further split into $n$ subsystems for each agent if the matrices $C,B_2$ in the output equation (\ref{eq:agent output}) have the same form to $A,B$ and $l=l_0n, o=o_0n$, i.e., $C = C_0\otimes I_n,B_2 = B_{2,0}\otimes I_n$ where $C_0\in\mathbb{R}^{l_0\times 3},B_{2,0}\in\mathbb{R}^{l_0\times o_0}$. Let us decompose 
the error $\pmb{e}(t)=\sum_{i=1}^{n} \pmb{e}_i(t)\otimes\mathbf{e}_i$, 
the control $u_{fb,i}(t)=\sum_{i=1}^{n} {u}_{fb,i}(t)\otimes\mathbf{e}_i$, 
the acuaor fault $f_{1,i}(t)=\sum_{i=1}^{n} {f}_{1,i}(t)\otimes\mathbf{e}_i$, 
the output $y(t)=\sum_{i=1}^{n} {y}_i(t)\otimes\mathbf{e}_i$, 
and the sensor fault $f_{2,i}(t)=\sum_{i=1}^{n} {f}_{2,i}(t)\otimes\mathbf{e}_i$ 
where $\pmb{e}_i(t)\in\mathbb{R}^3$, $u_{fb,i}(t)\in\mathbb{R}$, $f_{1,i}(t)\in\mathbb{R}^1$, $y_i(t)\in\mathbb{R}^{l_0}$, $f_{2,i}(t)\in\mathbb{R}^{o_0}$ and $\mathbf{e}_i$ is standard unit column vectors in $\mathbb{R}^n$, we get the $n$ subsystems:
\begin{equation} \label{eq:linear subsys}
    \begin{split}
        & \dot{\pmb{e}}_i(t)=A_0\pmb{e}_i(t)+B_0(u_{fb,i}(t)+f_{1,i}(t)) \\
        & {y}_i(t)=C_0\pmb{e}_i(t)+B_{2,0}f_{2,i}(t)   
    \end{split}
\end{equation}
\begin{remark}
    If the linearized system (\ref{eq:error}) can be splited into $n$ subsystems, this means that $\pmb{e}_i(t)$, the PID error of each state variable $x(t)$ of each agent in (\ref{eq:agent}), can be measured independently via sensors to obtain the independent outputs $y_i(t)$. In actual systems, this is usually done.
\end{remark}

By Theorem \ref{theorem1} again, the form shows that we can find the control gain $K_i\in\mathbb{R}^{1\times s}$ and observer gain $L_i\in\mathbb{R}^{s\times l_0}$ of the $i$th subsystem (\ref{eq:linear subsys}) that achieve the $H_\infty$ observer-based tracking control performance with a prescribed attenuation level $\rho_i$, where $s=3+(w_1+1)+(w_2+1)o_0$. The origin control gain $K$ of the origin agent system can be reconstructed by $K = \begin{bmatrix}
    \mathit{k}_1 & \mathit{k}_2 & ... & \mathit{k}_n
\end{bmatrix}^T, \mathit{k}_i=K_i^T\otimes\mathbf{e}_i$. The origin observer gain $L$ can be reconstructed in the same way.

@@

\textit{Agents:}\begin{enumerate}
    \item system parameters:
    Initial value $\tilde{x}(0) = [\pmb{e}(0)^T, F_1(0)^T\mathbin{,} F_2(0)^T, \tilde{e}(0)]^T = [[0.1X, ..., 0.1X]^T, 0, 0, 0]^T$ where $X \sim \mathcal{N}(0, 1)$. $C = C_0 \otimes I_n$ where $C_0 = I_3$.
    \item designed parameters: $\rho^*=30$. Actuator and sensor window size of smoothing signal model are $w_1=3$ and $w_2=4$, respectively.
\end{enumerate}

@@

    \item actuator coupling disturbance in (\ref{eq:UAV couple}):
    \\$c(t) = \sum_{k = 1, k \neq i}^{N_T}D_{i, 1, k}(x_{i, 1}(t))x_{k, 1}(t)\in\mathbb{R}^6$ where $D_{i, 1, k}(x_{i, 1}(t)) = diag(x_{i, 1, 1}(t)\mathbin{,}\dots\mathbin{,}x_{i, 1, 6}(t))$ with $x_{i, 1}(t) = [x_{i, 1, 1}(t)\mathbin{,}\dots\mathbin{,}x_{i, 1, 6}(t)]^T$.
    \item actuator non-coupling disturbance:
    \\ $d_1(t) = [100\sin(3t)\mathbin{,} ...\mathbin{,} 100\sin(3t)]^T\in\mathbb{R}^6$.
    \item sensor fault: $f_2(t)$ is set as a smoothed square wave as shown in Fig. \ref{fig:UAV, fa}.
\end{enumerate}
